Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-10-22 Methods-and-Results.tex
simple ML contact test adapted

To test the prerequisite effect of intergroup contact, every experience
sampling measurement started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (In person interaction for at least 10 minutes)}''.
Our participants recorded between 2--51 interactions with Dutch outgroup
members (3.23--91.07\% of individual experience sampling measurements;
31.59\% of all 1,225 experience sampling
responses)\footnote{Two participants only recorded two experience sampling measurements each and non of these included outgroup contacts. These participants are removed from any analyses that focus on outgroup contacts.}.

@@

We used a multilevel regression to check whether having an interaction
with an outgroup member had a situational (i.e., contemporaneous) effect
within the participants. We find that having an outgroup contact is
indeed associated with significantly more positive outgroup attitudes
within the participants (\textit{b} = 2.48, \textit{t}(1,200) = 4.37,
\textit{p} \textless{} .001, \textit{95\%CI}{[}1.37, 3.59{]}), even
after controlling for having an interaction with a non-Dutch (which did
not relate to outgroup attitudes independently; For full results see
Table \ref{tab:intergroupGeneralTblLong}, Figure
\ref{fig:ContactHypothesis}, and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.
While this offers support for the most basic premise of the intergroup
contact hypothesis (i.e., that any interaction will lead to more
positive outgroup attitudes), this analysis does not consider any notion
of interaction quality. We assess the conditional effects of interaction
numbers and interaction quality ratings after we report on the
individual studies to avoid under-powered analyses.

@@

To measure intergroup contacts, every experience sampling measurement
started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}''.
Participants were additionally offered a pop-up explanation: ``With
in-person interaction, we mean a continued interaction with another
person (potentially in a group) that lasted at least 10 minutes. This
interaction should be offline and face-to-face. It should include some
form of verbal communication and should be uninterrupted to still count
as the same interaction. Any individual interaction can last minutes or
hours. If there were multiple interaction partners, we would like you to
focus on the person that was most important to you during the
interaction.''. The participants recorded between 1--43 interactions
with Dutch majority people (1.64--93.75\% of individual experience
sampling measurements; 18.83\% of all 4,965 experience sampling
responses).

@@

We tested the most general contact hypothesis as in Study 1. We found a
significant correlation with average outgroup attitudes for both the
total number of outgroup interactions (\textit{r} = 0.22, \textit{p} =
0.019) and the number of measurement beeps with an interaction
(\textit{r} = 0.28, \textit{p} = 0.003). However, we found no
significant correlation between the participants' average interaction
quality and their average outgroup attitudes (\textit{r} = 0.04,
\textit{p} = 0.679). Thus, when considering the number of interactions
and average interaction quality jointly in a linear regression, we find
a main effect of intergroup contacts to predict outgroup attitudes
(\textit{b} = 0.81, \textit{t}(109) = 3.29, \textit{p} = 0.001,
\(\eta_p^2\) = 0.08) but no significant effect of average perceived
contact quality and no significant interaction term (for full results
see Table \ref{tab:intergroupGeneralTblLong} and Figure
\ref{fig:ContactHypothesis}). We additionally used a multilevel
regression to check whether having an interaction with an outgroup
member had a situational (i.e., contemporaneous) effect within the
participants. We find that having an outgroup interaction is indeed
associated with significantly more positive outgroup attitudes within
the participants (random slopes model; \textit{b} = 2.99,
\textit{t}(4,850) = 3.80, \textit{p} \textless{} .001,
\textit{95\%CI}{[}1.45, 4.54{]}), even after controlling for having an
interaction with a non-Dutch person (which did not relate to outgroup
attitudes independently; For full results see Table
\ref{tab:intergroupGeneralTblLong}, Figure \ref{fig:ContactHypothesis},
as well as Online Supplementary Material A). Thus, in our second data
set we find mixed results; outgroup contacts show a positive effect on
outgroup attitudes in the moment and on average between participants.
However, we find that this effect does not depend on the average
perceived interaction quality. This unexpected results, might be due to
the aggregation process and the following analyses will focus in on the
role of the perceived interaction quality during an outgroup
interaction.

@@

The measurement of intergroup contacts was identical to Study 2. The
participants recorded between 1--71 interactions with Dutch outgroup
members (2.13--87.18\% of individual experience sampling measurements;
41.44\% of all 4,107 experience sampling responses).

@@

As in Studies 1 and 2, we first conducted a general contact hypothesis
test. We found no correlation between the overall number of intergroup
contacts and average outgroup attitudes --- for neither the total number
of outgroup interactions nor the number of measurement beeps with an
interaction (\textbar{}\textit{r}\textbar{} \textless{} 0.13, \textit{p}
\textgreater{} 0.291). However, we did find a significant correlation
between the participants' average interaction quality and their average
outgroup attitudes (\textit{r} = 0.25, \textit{p} = 0.036). When
considering the number of interactions and average interaction quality
jointly in a linear regression, we only found a main effect of average
perceived interaction quality on outgroup attitudes (\textit{b} = 0.36,
\textit{t}(67) = 2.22, \textit{p} = 0.030, \(\eta_p^2\) = 0.07) but no
significant effect of the number of contacts nor an interaction effect.
This result mirrors that of Study 1 but is inconsistent with the second
study. Additionally, in a multilevel regression to check for
contemporaneous within person effects, we found that having an outgroup
interaction was indeed associated with significantly more positive
outgroup attitudes within the participants (random slopes model;
\textit{b} = 5.63, \textit{t}(3,834) = 6.61, \textit{p} \textless{}
.001, \textit{95\%CI}{[}3.97, 7.29{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently; for full results see Table
\ref{tab:intergroupGeneralTblLong}, Figure \ref{fig:ContactHypothesis},
and Online Supplementary Material A). Thus, in our third data set we
again found mixed results. Aggregate outgroup contacts were generally
not significantly related to average outgroup attitudes on a between
participant level and this effect was also not suppressed by an average
interaction quality (i.e., no interaction effect). However, while the
aggregate data was partly inconsistent with the first two studies, the
within person contemporaneous effect of intergroup contact was
consistent across all three studies.

>>>>

To test the prerequisite effect of intergroup contact, every experience
sampling measurement started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (In person interaction for at least 10 minutes)}''.
Our participants recorded between 2--51 interactions with Dutch outgroup
members (3.23--91.07\% of individual experience sampling measurements;
387 of all 1,225 experience sampling
responses)\footnote{Two participants only recorded two experience sampling measurements each and non of these included outgroup contacts. These participants are removed from any analyses that focus on outgroup contacts.}.

@@

We used a multilevel regression to check whether having an interaction
with an outgroup member had a situational (i.e., contemporaneous) effect
within the participants. We find that having an outgroup contact is
indeed associated with significantly more positive outgroup attitudes
within the participants (\textit{b} = 2.48, \textit{t}(1,200) = 4.37,
\textit{p} \textless{} .001, \textit{95\%CI}{[}1.37, 3.59{]}), even
after controlling for having an interaction with a non-Dutch (which did
not relate to outgroup attitudes independently). Additionally, while
multi-level regressions are generally robust against unequal cell sizes,
we correct for inequalities by use centered predictors and reintroducing
the means as level two predictors. Nonetheless, we also offer an
analysis of the joint data sets in Online Supplemental Material A (for
full results see Table \ref{tab:intergroupGeneralTblLong}, Figure
\ref{fig:ContactHypothesis}, and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.
While this offers support for the most basic premise of the intergroup
contact hypothesis (i.e., that any interaction will lead to more
positive outgroup attitudes), this analysis does not consider any notion
of interaction quality. We assess the conditional effects of interaction
numbers and interaction quality ratings after we report on the
individual studies to avoid under-powered analyses.

@@

To measure intergroup contacts, every experience sampling measurement
started with the question
``\textit{Did you meet a Dutch person this morning [/afternoon]? (in-person interaction for at least 10 minutes)}''.
Participants were additionally offered a pop-up explanation: ``With
in-person interaction, we mean a continued interaction with another
person (potentially in a group) that lasted at least 10 minutes. This
interaction should be offline and face-to-face. It should include some
form of verbal communication and should be uninterrupted to still count
as the same interaction. Any individual interaction can last minutes or
hours. If there were multiple interaction partners, we would like you to
focus on the person that was most important to you during the
interaction.''. The participants recorded between 1--43 interactions
with Dutch majority people (1.64--93.75\% of individual experience
sampling measurements; 935 of all 4,965 experience sampling responses).

@@

We tested the most general contact hypothesis as we did for Study 1. We,
again, used a multilevel regression analysis to check whether having an
interaction with an outgroup member had a situational (i.e.,
contemporaneous) effect within the participants. We find that having an
outgroup interaction is indeed associated with significantly more
positive outgroup attitudes within the participants (random slopes
model; \textit{b} = 2.83, \textit{t}(4,850) = 3.57, \textit{p}
\textless{} .001, \textit{95\%CI}{[}1.28, 4.38{]}), even after
controlling for having an interaction with a non-Dutch person (which did
not relate to outgroup attitudes independently). We again added the
participant means back into the model. Because these are the means of
the contact dummy variables they are the proportion of measurements that
included a particular contact. This not only adjusts the model for
unequal cell sizes but allows us to see whether between subject
differences in the relative number of interactions had an effect on
outgroup attitudes. We find that in this data set participant-level
outgroup contact proportions were also a positive predictor of outgroup
attitudes (\textit{b} = 26.55, \textit{t}(110) = 2.90, \textit{p} =
0.004, \textit{95\%CI}{[}8.61, 44.46{]}). The relative number of
non-outgroup interactions showed no such effect (for full results see
Table \ref{tab:intergroupGeneralTblLong}, Figure
\ref{fig:ContactHypothesis}, as well as Online Supplementary Material
A). Thus, in our second data set we find that outgroup contacts show a
positive effect on outgroup attitudes in the moment and on average
between participants.

@@

The measurement of intergroup contacts was identical to Study 2. The
participants recorded between 1--71 interactions with Dutch outgroup
members (2.13--87.18\% of individual experience sampling measurements;
1,702 of all 4,107 experience sampling responses).

@@

Additionally, in a multilevel regression to check for contemporaneous
within person effects, we found that having an outgroup interaction was
indeed associated with significantly more positive outgroup attitudes
within the participants (random slopes model; \textit{b} = 5.57,
\textit{t}(3,834) = 6.52, \textit{p} \textless{} .001,
\textit{95\%CI}{[}3.90, 7.23{]}), even after controlling for having an
interaction with a non-Dutch (which did not relate to outgroup attitudes
independently; for full results see Table
\ref{tab:intergroupGeneralTblLong}, Figure \ref{fig:ContactHypothesis},
and Online Supplementary Material A). Thus, in our third data set we
found within person contemporaneous effect of intergroup contact was
consistent across all three studies.
